---
title: "Butt et al: scRNAseq analysis - Final analyses"
author: "Agnes A. Steixner-Kumar"
date: "20 11 2020"
output: 
  pdf_document:
    latex_engine: xelatex 
    fig_caption: yes
    fig_height: 8
    fig_width: 13
    keep_tex: no
    number_sections: yes
    toc: yes
  html_notebook: 
    fig_caption: yes
    fig_height: 8
    fig_width: 13
    number_sections: yes
    toc: yes
    df_print: paged
  html_document: 
    fig_caption: yes
    fig_height: 8
    fig_width: 13
    number_sections: yes
    toc: yes
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, echo=FALSE, warning=F}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      echo = TRUE)
```

```{r packrat, echo=FALSE, eval=F}
options(repos=structure(BiocManager::repositories()))
packrat::init()
packrat::snapshot()
```

# Load libraries and data
## Load libraries
```{r load_lib, paged.print=TRUE, cache=FALSE}
library(reshape2)
library(RColorBrewer)
library(grid)
library(Seurat) #v3.0.0
library(cowplot)
library(dplyr)
library(ggplot2)
library(scales)
```

## Load data
This is the final .RDS object as uploaded on GEO (GSExxx)
```{r load_data, paged.print=TRUE}
ODD<-readRDS("F:/GEO_submission_Butt_et_al/Butt_et_al_hypoxia_Seurat.RDS")
```

# Analysis
## Figure 4A
### Define colors
```{r config_colors, paged.print=TRUE}
identities <- levels(Idents(ODD))
my_color_palette <- hue_pal()(length(identities))
my_color_palette <-c('brown2', my_color_palette[2:16])
```

### Shift ependymal cluster upwards
```{r shift_embedding, paged.print=TRUE}
emb<-Embeddings(ODD, reduction = "umap")
#shift selected cells upwards on UMAP2
emb[,2][emb[,2]<(-15)]<-emb[,2][emb[,2]<(-15)]+12
#feed changed embeddings back into object
ODD@reductions$umap@cell.embeddings<-emb
```

### Make plot of UMAP embedding
```{r dimplot, fig.align="center", paged.print=TRUE}
DimPlot(ODD, reduction = 'umap', label =F, repel = T, cols = my_color_palette)+NoLegend()+NoAxes()
```

## Figure 4B
### Create dataframe of ODD &/or tdTomato+ cells
```{r ODD_tdT_df, paged.print=TRUE}
# prepare dataframe
ODD_tdTom_cluster<-data.frame(matrix(nrow = ncol(ODD), ncol = 5), row.names = colnames(ODD))
colnames(ODD_tdTom_cluster)=c("Cluster", "Group", "ODD", "tdTomato", "Expression")
# fill in metadata
ODD_tdTom_cluster$Cluster<-Idents(ODD)
ODD_tdTom_cluster$Sample<-ODD$Sample
ODD_tdTom_cluster$Group<-ODD$group
# get expression values
ODD_tdTom_cluster$ODD<-GetAssayData(ODD)["ODD",]
ODD_tdTom_cluster$tdTomato<-GetAssayData(ODD)["tdTomato",]
# recode
ODD_tdTom_cluster$Expression[ODD_tdTom_cluster$ODD==0 & ODD_tdTom_cluster$tdTomato==0]<-"none"
ODD_tdTom_cluster$Expression[ODD_tdTom_cluster$ODD>0 & ODD_tdTom_cluster$tdTomato==0]<-"ODD"
ODD_tdTom_cluster$Expression[ODD_tdTom_cluster$ODD==0 & ODD_tdTom_cluster$tdTomato>0]<-"tdTomato"
ODD_tdTom_cluster$Expression[ODD_tdTom_cluster$ODD>0 & ODD_tdTom_cluster$tdTomato>0]<-"both"
```
### Count positive cells
```{r count_ODD_tdT, paged.print=TRUE}
# create count table
ODD_tdTom_cluster_counts<-ODD_tdTom_cluster %>%
 group_by(Cluster, Expression, Group) %>%
    count()
# order levels for display
ODD_tdTom_cluster_counts$Expression<-factor(ODD_tdTom_cluster_counts$Expression, levels=c("none", "both", "tdTomato", "ODD"))
```

### Clean up cluster lables
```{r clean_labels, paged.print=TRUE}
cluster_labels<-as.character(unique(ODD_tdTom_cluster_counts$Cluster))
cluster_labels<-gsub('_', ' ', cluster_labels)
cluster_labels<-gsub('cells', '', cluster_labels)
cluster_labels<-gsub(' $', '', cluster_labels)
cluster_labels
```
### Plot proportions
```{r ODD_tdT_prop_graph, fig.align="center", paged.print=TRUE}
ggplot(ODD_tdTom_cluster_counts, aes(y=n, x=Cluster, fill=Expression)) +  
  geom_bar( stat="identity", position="fill")+
  scale_y_continuous(labels= scales::percent)+
  ylab("Construct expressing cells")+ 
  coord_cartesian(expand = FALSE)+
  scale_fill_manual(values=c("gray72", "lightgoldenrod1", "lightcoral", "darkseagreen3"))+ 
  annotate(geom = 'text',size=5, label=cluster_labels, y=0.99, x=seq(0:15), angle=90, hjust=1)+
  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank(), 
        axis.text.y=element_text(size=19), axis.title=element_text(size=20), legend.text=element_text(size=18), 
        legend.title=element_text(size=20),legend.position = "bottom")+
  theme(plot.margin=unit(c(0.3,3.6,0,0.7), "cm"))
```

## Figure 4C
### Split objects by group
```{r split_nor_hyp, paged.print=TRUE}
Idents(ODD)<-ODD$group
ODD_norm<-subset(ODD, idents = "Normoxia")
ODD_hyp<-subset(ODD, idents = "Hypoxia")
```

### Make normoxia feature plots
```{r normoxia_ftplots, fig.align="center", paged.print=TRUE}
lapply(c("ODD", "tdTomato", "Vegfa"), function(x) FeaturePlot(
            ODD_norm, features = x, pt.size = 0.7, order=F, min.cutoff = 0.5,
            max.cutoff = 2, cols = (c("bisque2", "black")))+
            NoAxes()+
            ggtitle(x)+
            NoLegend())
# order cells with regard to expression to prevent masking
FeaturePlot(ODD_norm, features = "Hk2", pt.size = 0.7, order=T, min.cutoff = 0.5,
            max.cutoff = 2, cols = c("bisque2", "black"))+
            NoAxes()+
            ggtitle("Hk2")+
            NoLegend()
```

### Make hypoxia feature plots
```{r hypoxia_ftplots, fig.align="center", paged.print=TRUE}
lapply(c("ODD", "tdTomato", "Vegfa"), function(x) FeaturePlot(
            ODD_hyp, features = x, pt.size = 0.7, order=F, min.cutoff = 0.5,
            max.cutoff = 2, cols = (c("bisque2", "black")))+
            NoAxes()+
            ggtitle(x)+
            NoLegend())
# order cells with regard to expression to prevent masking
FeaturePlot(ODD_hyp, features = "Hk2", pt.size = 0.7, order=T, min.cutoff = 0.5,
            max.cutoff = 2, cols = c("bisque2", "black"))+
            NoAxes()+
            ggtitle("Hk2")+
            NoLegend()
```
## Figure 4D
### Create rough identity clusters
These cell types correspond to cell types distinguished in immunofluorescence quantifications
```{r rough_clusters, paged.print=TRUE}
ODD$rough_ident<-NA
# merge all neuronal types
ODD$rough_ident[ODD$final_identity=="Glutamatergic1"|ODD$final_identity=="Glutamatergic0"|ODD$final_identity=="Glutamatergic2"|
               ODD$final_identity=="Glutamatergic3"|ODD$final_identity=="Glutamatergic4"|ODD$final_identity=="Gabaergic"|ODD$final_identity=="Mossy_cells"]<-"Neurons"
ODD$rough_ident[ODD$final_identity=="Astrocytes"]<-"Astrocytes"
ODD$rough_ident[ODD$final_identity=="Oligodendrocytes"|ODD$final_identity=="OPC"]<-"OPC, Oligo"
ODD$rough_ident[ODD$final_identity=="Microglia"]<-"Microglia"
ODD$rough_ident[ODD$final_identity=="Endothelial"]<-"Endothelial"
```
### Create dataframe of binary Hk2 expression 
```{r Hk2_df, paged.print=TRUE}
# prepare dataframe
Hk2_cluster<-data.frame(matrix(nrow = sum(!is.na(ODD$rough_ident)), ncol = 4), row.names = colnames(ODD)[!is.na(ODD$rough_ident)])
colnames(Hk2_cluster)=c("Cluster", "Group", "Hk2", "Hif1a")
# fill in metadata
Hk2_cluster$Cluster<-ODD$rough_ident[!is.na(ODD$rough_ident)]
Hk2_cluster$Sample<-ODD$Sample[!is.na(ODD$rough_ident)]
Hk2_cluster$Group<-ODD$group[!is.na(ODD$rough_ident)]
# fill in expression data and recode
Hk2_cluster$Hk2<-GetAssayData(ODD)["Hk2",!is.na(ODD$rough_ident)]
Hk2_cluster$Hk2_yn[Hk2_cluster$Hk2>0]<-"yes"
Hk2_cluster$Hk2_yn[Hk2_cluster$Hk2==0]<-"no"
# create count table
Hk2_cluster_prop<-Hk2_cluster %>%
 group_by(Cluster, Group, Hk2_yn) %>%
    summarise(n=n()) %>%
        mutate(Percent=(n/sum(n))*100)

```
### Perform chi-square tests per cluster
```{r chisq_cluster, paged.print=TRUE}
lapply(unique(Hk2_cluster_prop$Cluster), function(x) {
  paste0(x, ": ",
  chisq.test(data.frame(c(Hk2_cluster_prop$n[Hk2_cluster_prop$Cluster==x][1],
                          Hk2_cluster_prop$n[Hk2_cluster_prop$Cluster==x][2]),
                        c(Hk2_cluster_prop$n[Hk2_cluster_prop$Cluster==x][3],
                          Hk2_cluster_prop$n[Hk2_cluster_prop$Cluster==x][4])), 
              correct = T)$p.value)
  }
       )
```
### Perform chi-square tests for total sample
```{r chisq_total, paged.print=TRUE}
# count positive/negative cells per group
Hk2_cluster_prop_total<-Hk2_cluster %>%
 group_by(Group, Hk2_yn) %>%
    summarise(n=n()) %>%
        mutate(Percent=(n/sum(n))*100)
# perform chi-square test
chisq.test(data.frame(c(Hk2_cluster_prop_total$n[1],
                        Hk2_cluster_prop_total$n[2]),
                      c(Hk2_cluster_prop_total$n[3],
                        Hk2_cluster_prop_total$n[4])), correct = F)$p.value
```
### Plot Hk2+ proportions per cluster
```{r Hk2_prop_plot, fig.align="center", paged.print=TRUE}
# make factor, order levels
Hk2_cluster_prop$Cluster<-factor(Hk2_cluster_prop$Cluster, levels=c("Endothelial", 'Neurons', 'Astrocytes', "OPC, Oligo", 'Microglia'))

# plot
ggplot(data = Hk2_cluster_prop[Hk2_cluster_prop$Hk2_yn=="yes",], aes(x = Cluster, y = Percent)) +
  geom_col(aes(fill = Group, width=0.6), col='black' , show.legend = FALSE) +
  coord_flip()+ facet_wrap(~Group)+
  theme(axis.ticks.y=element_blank(), axis.text.y=element_text(size=22, angle = 0), 
        axis.text.x=element_text(size=25), axis.title.y=element_blank(), axis.title.x=element_text(size=25), 
        legend.text=element_text(size=16), legend.title=element_text(size=25))+
  scale_fill_manual(values = c("#FFBDBD","#512274"))+
  theme(strip.text.x = element_text(size = 30, colour = "black",margin = margin(0.5,0,0.5,0, "cm")))+
  theme(aspect.ratio = 1/1.1)+
  scale_y_continuous( limits = c(0,20), expand = c(0,0))+
  theme(panel.spacing = unit(2, "lines"))+
  ylab('Percentage of Hk2 positive cells')+
  theme(plot.margin=unit(c(0,0.5,0,0), "cm"))
```

## Differential expression testing
### Vegfa 
```{r DE_Vegfa, paged.print=TRUE}
# Test for overall differential expression between Normoxia and Hypoxia
FindMarkers(ODD, ident.1 = 'Hypoxia', ident.2 = 'Normoxia', group.by = 'group', features = 'Vegfa')
# Test for differential expression among clusters
Idents(ODD)<-ODD$final_identity
FindAllMarkers(ODD, features = "Vegfa")
```
### Hk2
```{r DE_Hk2, paged.print=TRUE, cache=FALSE}
Idents(ODD)<-ODD$group
# Microglial expression versus all other cells under normoxia
FindMarkers(ODD, ident.1 = 'Microglia', subset.ident = 'Normoxia', group.by = 'final_identity', features = 'Hk2', logfc.threshold = 0.01)
# Microglial expression versus all other cells under hypoxia
FindMarkers(ODD, ident.1 = 'Microglia', subset.ident = 'Hypoxia', group.by = 'final_identity', features = 'Hk2', logfc.threshold = 0.01)
```
## Supplementary info
### Suppl. Figure 3
```{r ODD_tdT_vlnplots, fig.align="center", paged.print=TRUE, cache=FALSE}
Idents(ODD)<-ODD$final_identity
# tdTomato violin plots
VlnPlot(ODD, features = "tdTomato", cols = my_color_palette, pt.size = 0.1)+
  NoLegend()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  theme(plot.title = element_text(hjust = 0.5))
# ODD violin plots
VlnPlot(ODD, features = "ODD", cols = my_color_palette, pt.size = 0.1)+
  NoLegend()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  theme(plot.title = element_text(hjust = 0.5))
```
Test for expression differences in tdTomato and ODD between microglia and neurons
```{r DE_ODD_tdT, paged.print=TRUE}
FindMarkers(ODD, features = c("tdTomato", "ODD"), logfc.threshold = 0.01, ident.1 = "Microglia", ident.2 = c("Glutamatergic0", "Glutamatergic1", "Glutamatergic2", "Glutamatergic3", "Glutamatergic4", "Gabaergic", "Mossy_cells"))
```
### Suppl. Figure 4A
```{r vegfa_vlnplots, fig.align="center", paged.print=TRUE}
# Vegfa violin plots
VlnPlot(ODD, features = "Vegfa", cols = c("pink","purple"), pt.size = 0.1, split.by = "group")+
  NoLegend()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  theme(plot.title = element_text(hjust = 0.5))
```
### Suppl. Figure 4B
```{r DE_vegfa, paged.print=TRUE}
pvals_Vegfa<-lapply(levels(ODD), function(x) FindMarkers(ODD, ident.1 = "Hypoxia", ident.2 = "Normoxia", subset.ident = x, group.by = "group", features = "Vegfa", logfc.threshold = 0.01))

names(pvals_Vegfa)<-levels(ODD)
dplyr::bind_rows(pvals_Vegfa, .id = "Cluster")
```
## Session info
```{r session_info, tidy=TRUE, tidy.opts=list(width.cutoff=80)}
sessionInfo()
```